---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
```

# bootUR
`bootUR` implements several bootstrap tests for unit, both for single time series and for (potentially) large systems of time series.

## Installation
`bootUR` can be installed using
```{r, eval = FALSE}
devtools::install_github("smeekes/bootUR")
```

## Inspect data for missing values
The bootstrap tests in `bootUR` do not work with missing data, although multivariate time series with different start and end dates (unbalanced panels) are allowed. `bootUR` provides a simple function to check if your data contain missing values. We will illustrate this on the `MacroTS` dataset of macreconomic time series that comes with the package.

```{r}
library(bootUR)
data("MacroTS")
check_missing_insample_values(MacroTS)
```

## Checking start and end points of time series
If your time series have different starting and end points (and thus some series contain `NA`s at the beginning and/or end of your sample, the resampling-based moving block bootstrap (MBB) and sieve bootstrap (SB) cannot be used. `bootUR` lets you check the start and end points as follows:

```{r}
sample_check <- find_nonmissing_subsample(MacroTS)
sample_check$range  # Provides the number of the first and last non-missing observation for each series
sample_check$all_equal  # Gives TRUE if the time series all start and end at the same observation
```

## Univariate Dickey-Fuller unit root test
To perform a standard augmented Dickey-Fuller (ADF) bootstrap unit root test on a single time series, use the `boot_df()` function. The function allows to set many options, including the bootstrap method used (option `boot`), the deterministic components included (option `dc`) and the type of detrending used. While `dc = "OLS"` gives the standard ADF test, `dc = "QD"` provides the powerful DF-GLS test of Elliott, Rothenberg and Stock (1996). Here we use the terminology Quasi-Differencing (QD) rather than GLS as this conveys the meaning less ambiguously and is the same terminology used by Smeekes and Taylor (2012) and Smeekes (2013).

We illustrate the bootstrap ADF test here on Dutch GDP, with the sieve bootstrap (`boot = SB`) to as the specification used by Palm, Smeekes and Urbain (2008) and Smeekes (2013). We set only 399 bootstrap replications (`B = 399`) to prevent the code from running too long. We add an intercept and a trend (`dc = 2`), and compare OLS with QD(GLS) detrending. The option `verbose = TRUE` prints easy to read output on the console.

```{r}
GDP_NL <- MacroTS[, 4]
adf_out <- boot_df(GDP_NL, B = 399, boot = "SB", dc = 2, detr = c("OLS", "QD"), verbose = TRUE)
```

Use `boot_union()` for a test based on the union of rejections of 4 tests with different number of deterministic components and different type of detrending (Smeekes and Taylor, 2012). The advantage of the union test is that you don't have to specify these (rather influential) specification tests. This makes the union test a safe option for quick or automatic unit root testing where careful manual specification setup is not viable. Here we illustrate it with the sieve wild bootstrap as proposed by Smeekes and Taylor (2012).

```{r}
union_out <- boot_union(GDP_NL, B = 399, boot = "SWB", verbose = TRUE)
```

## Panel Unit Root Test
The function `panel_test` performs a test on a multivariate (panel) time series by testing the null hypothesis that all series have a unit root. A rejection is typically interpreted as evidence that a 'significant proportion' of the series is stationary, although how large that proportion is - or which series are stationary - is not given by the test. The test is based on averaging the individual test statistics, also called the Group-Mean (GM) test in Palm, Smeekes and Urbain (2011).

Palm, Smeekes and Urbain (2011) introduced this test with the moving block bootstrap (`boot = "MBB"`), which is the standard option. However, this resampling-based method cannot handle unbalancedness, and will therefore gives an error when applied to `MacroTS`. Therefore, you should switch to one of the wild bootstrap methods. Here we illustrate it with the dependent wild bootstrap (DWB) of Shao (2010).

By default the union test is used for each series (`union = TRUE`), if this is set to `FALSE` the deterministic components and detrending methods can be specified as in the univariate Dickey-Fuller test.

Although the sieve bootstrap method `"SB"` and `"SWB"` can be used (historically they have been popular among practitioners), Smeekes and Urbain (2014) show that these are not suited to capture general forms of dependence across units. The code will give a warning to recommend using a different bootstrap method.

```{r}
# This will give an error!
# panel_out <- paneltest(MacroTS, boot = "MBB", B = 399, verbose = TRUE)
panel_out <- paneltest(MacroTS, boot = "DWB", B = 399, verbose = TRUE)
```

## Individual ADF Tests for Multiple Time Series
To perform individual ADF tests on multiple time series simultaneously, the function `iADFtest()` can be used. As the bootstrap is performed for all series simultaneously, resampling-based bootstrap methods `"MBB"` and `"SB"` cannot be used directly in case of unbalanced panels. If they are used anyway, the function will revert to splitting the bootstrap up and perfoming it individually per time series. In this case a warning is given to alert the user. The other options are the same as for `paneltest`.
```{r}
iADF_out <- iADFtest(MacroTS[, 1:5], boot = "MBB", B = 399, verbose = TRUE, union = FALSE, dc = 2, detr = "OLS")
```

## Sequential Tests for Multiple Time Series
Smeekes (2015). Bootstrap Sequential Quantile Test (BSQT).
The other options are the same as for `paneltest`.
```{r}
BSQT_out <- BSQTtest(MacroTS[, 16:20], boot = "AWB", B = 399, verbose = TRUE)
```

## References
* Elliott, Rothenberg and Stock (1996)
* Palm, Smeekes and Urbain (2011)
* Shao (2010)
* Smeekes (2013)
* Smeekes and Urbain (2014)
* Smeekes and Taylor (2012)
