---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
```

# bootUR: Bootstrap Unit Root Tests
<!-- badges: start -->
[![CRAN_Version_Badge](https://www.r-pkg.org/badges/version/bootUR)](https://cran.r-project.org/package=bootUR)
[![CRAN_Downloads_Badge](https://cranlogs.r-pkg.org/badges/grand-total/bootUR)](https://cran.r-project.org/package=bootUR)
[![License_GPLv2_Badge](https://img.shields.io/badge/License-GPLv2-yellow.svg)](https://www.gnu.org/licenses/old-licenses/gpl-2.0.html)
[![License_GPLv3_Badge](https://img.shields.io/badge/License-GPLv3-yellow.svg)](https://www.gnu.org/licenses/gpl-3.0.html)
[![R-CMD-check](https://github.com/smeekes/bootUR/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/smeekes/bootUR/actions/workflows/R-CMD-check.yaml)
<!-- badges: end -->

The R package `bootUR` implements several bootstrap tests for unit roots, both for single time series and for (potentially) large systems of time series.

## Installation and Loading
The package can be installed from CRAN using
```{r install-cran, eval = FALSE}
install.packages("bootUR")
```

The development version of the `bootUR` package can be installed from GitHub using
```{r install-github, eval = FALSE}
# install.packages("devtools")
devtools::install_github("smeekes/bootUR")
```
When installing from GitHub, in order to build the package from source, you need to have the appropriate R development tools installed ([Rtools](https://cran.r-project.org/bin/windows/Rtools/) on Windows, or [these tools](https://mac.r-project.org/tools/) on Mac).

If you want the vignette to appear in your package when installing from GitHub, use
```{r install-vign, eval = FALSE}
# install.packages("devtools")
devtools::install_github("smeekes/bootUR", build_vignettes = TRUE, dependencies = TRUE)
```
instead. As building the vignette may take a bit of time (all bootstrap code below is run), package installation will be slower this way.

After installation, the package can be loaded in the standard way:
```{r load}
library(bootUR)
```

## Functionality
A quick overview of the package functionality is provided in the vignette `vignette("bootUR")`. 

A further investigation of the functionalities is provided in the *Journal of Statistical Software* article [bootUR: An R Package for Bootstrap Unit Root Tests ](https://doi.org/10.18637/jss.v106.i12).
